### 2021-07-22 appendix/distributor-first-floor.tex

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/rozdelovac-prizemi/rozdelovac-prizemi-celni-strana.png}
    \caption{Rozdělovač přízemí – čelní strana.}
    \label{fig:rozdelovac-prizemi-celni-strana}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/rozdelovac-prizemi/rozdelovac-prizemi.png}
    \caption{Rozdělovač přízemí – instalace uvnitř.}
    \label{fig:rozdelovac-prizemi}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/rozdelovac-prizemi/rozdelovac-prizemi-umisteni-dps.png}
    \caption{Rozdělovač přízemí – umístění DPS}
    \label{fig:rozdelovac-prizemi-umisteni-dps}
\end{figure}