### 2021-09-17 appendix/i2c-hub.tex
Change text and images.

\section{Rozbočovač I$^2$C}
\label{app:rozbocovac-i2c}
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/krb/rozbocovac-i2c-spodni-strana.png}
    \caption{Rozbočovač I$^2$C – spodní strana DPS.}
    \label{fig:rozbocovac-i2c-spodni-strana}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/krb/rozbocovac-i2c-vrchni-strana.png}
    \caption{Rozbočovač I$^2$C – vrchní strana DPS.}
    \label{fig:rozbocovac-i2c-vrchni-strana}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/krb/rozbocovac-i2c-celni-strana.png}
    \caption{Rozbočovač I$^2$C – čelní strana DPS.}
    \label{fig:rozbocovac-i2c-celni-strana}
\end{figure}